\pkg{sdmTMB}: An \proglang{R} Package for Fast, Flexible, and User-Friendly Generalized Linear Mixed Effects Models with Spatial and Spatiotemporal Random Fields

### 2023-09-23 doc/paper-jss/sdmTMB-paper.Rnw
Edits

This paper describes the underlying statistical models (Section~\ref{model-description}), explains how \pkg{sdmTMB} is designed and summarizes its software functionality (Section~\ref{sec:software}), illustrates its use through three case studies (Sections~\ref{pcod}, \ref{dogfish}, and \ref{svc-owls}), compares \pkg{sdmTMB} to other packages (Section~\ref{package-comparisons}) and concludes with a discussion and comparison to related software (Section~\ref{discussion}).

@@

\pkg{sdmTMB} would not be possible without the \pkg{TMB} \citep{kristensen2016}
and \proglang{R}-\pkg{INLA} \citep{rue2009, lindgren2011, lindgren2015}
\proglang{R} packages. \pkg{sdmTMB} is heavily inspired by and in some places
code has been adapted from both the \pkg{VAST} \citep{thorson2019} and
\pkg{glmmTMB} \citep{brooks2017} \proglang{R} packages (as described in the
DESCRIPTION and inst/COPYRIGHTS files). Smoother support was possible thanks to
\pkg{mgcv} \citep{wood2017a}. We thank the authors of all these packages. We
thank S. Kotwicki, M. Lindmark, M. Martin, C.C. Monnahan, P.M. Regular, J.T.
Thorson, and J. Watson for helpful comments that substantially improved the
manuscript. Christmas Bird Count data were provided by National Audubon Society
and through the generous efforts of Bird Studies Canada and countless volunteers
across the Western Hemisphere.

>>>>

This paper describes the underlying statistical models (Section~\ref{model-description}), explains how \pkg{sdmTMB} is designed and summarizes its software functionality (Section~\ref{sec:software}), illustrates its use through three case studies (Sections~\ref{pcod}, \ref{dogfish}, and \ref{svc-owls}), compares \pkg{sdmTMB} to other packages (Section~\ref{package-comparisons}) and concludes with a discussion of TODO  (Section~\ref{discussion}).

@@

\pkg{sdmTMB} would not be possible without the \pkg{TMB} \citep{kristensen2016} and \proglang{R}-\pkg{INLA} \citep{rue2009, lindgren2011, lindgren2015} \proglang{R} packages. 
\pkg{sdmTMB} is heavily inspired by and in some places code has been adapted from both the \pkg{VAST} \citep{thorson2019} and \pkg{glmmTMB} \citep{brooks2017} \proglang{R} packages (as described in the \texttt{DESCRIPTION} and \texttt{inst/COPYRIGHTS} files). 
Smoother support was possible thanks to \pkg{mgcv} \citep{wood2017a}. 
We thank the authors of all these packages. 
We thank S. Kotwicki, M. Lindmark, M. Martin, C.C. Monnahan, P.M. Regular, J.T. Thorson, J. Watson, two anonymous reviewers, and the editor for helpful comments that substantially improved the manuscript. 
Christmas Bird Count data were provided by National Audubon Society and through the generous efforts of Bird Studies Canada and countless volunteers across the Western Hemisphere.
